DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-06-27 2015CompleteDerivation/completeDerivation.tex
Adding justification for differentiability

	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbbm{R}$. $\forall \mathcal{c} \in \mathcal{C} \exists ! \rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$. As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$. $\mathcal{L}(S,\mu) = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}. Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$. $\varrho$ is a homomorphism as we saw in the proof for \ref{classical_vector space}. $\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
\end{justification}

@@

Moreover, the density itself has to be differentiable. The distribution $\rho_\mathcal{c}$ is a real value function of the state. As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$. As a state variable it identifies the states by the density of the material associated with them. This is more physically meaningful as one may first suspect, since placing physical markers is a common way to define the references for a coordinate system, which in the end is placing material at particular states. In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differential as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$. As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist. Therefore the density itself is differentiable.

\begin{prop}\label{differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a homogeneous classical material identified by a finite number of continuous independent state variables is a differentiable manifold. The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	TODO
%Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$. Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function. If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$. There exists an outcome in $\mathcal{S}$ that can physically distinguish the two distributions. There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho(\mathcal{c}_1)=\rho$. $\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective. $\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
\end{justification}

As we identified the space of Lebesque integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill. Some are not restrictive enough. The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical. The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ not only include non-differentiable functions, but also include functions whose integral is infinite, which would represent infinite amount of material. Such functions are also present in the space of differentiable functions $C^1(\mathcal{S})$. Some definitions are too restrictive. Requiring compact support (i.e.the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states. Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range. Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible. }

@@

If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary. This means we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This means that we cannot change physical units of all the variables as we like: they are part of a unit system. As we know, changing units of position will also change units of momentum. Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units. An element of $\mathcal{S}$ will be a geometrical object in such space. For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined. As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$. As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant. They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant. They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.\footnote{The components $k$ are the classical analogue of the wave numbers.} In other words: a covector is the simplest local geometrical object that allows coordinate independent densities. Physically, the component $k_i$ of a covector, for example the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$. The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation.  We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.

If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$. These variables are independent, so the total number of possibilities in a small region of $\mathcal{S}$ is proportional to $dq \wedge dk$. That is if we double the range of $dq$ or $dk$ then the number of possibilities doubles. We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$. That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property. Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait. That is why one can choose "natural units" for which $\hbar=1$. In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e. compact) region $U \subseteq \mathcal{S}$ will be given by $\int_U \hbar dq \wedge dk$.

If $\mathcal{Q}$ is $n$-dimensional, $d\mu$ is proportional to $dq^n \wedge dk_n$. But it's not just the volume that is preserved as we change state variables. Each pair $(q^i, k_i)$ represents an independent set of state variables, with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$. We call such a pair a degree of freedom. As they are independent the total number of states is the product of the possibilities of each d.o.f.: $\hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$. This is equivalent to saying that independent d.o.f. are orthogonal surfaces in $\mathcal{S}$.

@@

\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distributions associated with a composite state $\mathcal{c} \in \mathcal{C}$. $\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$. Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form. The description of the simplest physical object on which we can define a state-variable-invariant must be given by at least one continuous state variable $q$. Let $\mathcal{Q}$ be the manifold charted by $q$. $\mathcal{Q}$ is an embedded manifold of $\mathcal{S}$. $\mathcal{Q}$ is differentiable. Let $q$ fully define the unit system for the description of the physical object. Then the physical object is represented by a geometrical object on $\mathcal{Q}$. The physical object is infinitesimal, therefore is represented by a local geometrical object. The simplest geometrical object in $\mathcal{Q}$ is a point, charted by $q$. A density over points is not invariant: $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$. Next we consider local geometrical objects charted by 2 variables. Densities over vectors $v(q) e_q \in T\mathcal{Q}$ are not invariant: $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$. Densities over covectors $k(q) dq \in T^*\mathcal{Q}$ are invariant: $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. Covectors are the simplest geometrical object to allow unit independent densities. The state space $\mathcal{S}$ for the simplest object that allows state-variable-independent densities is $T^*\mathcal{Q}$ of a one dimensional differential manifold $\mathcal{Q}$. Express integration in coordinates. $\int_{\mathcal{S}} \rho_\mathcal{c} d\mu \propto \int_{\mathcal{S}} \rho_\mathcal{c}(q, k) dq \wedge dk$. $d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant, $p\equiv \hbar k$  and $\omega$ is the canonical two form. $\omega$ is invariant under state variable change. $\omega$ is a symplectic metric for $\mathcal{S}=T^*\mathcal{Q}$. $\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.

>>>>

	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbbm{R}$. $\forall \mathcal{c} \in \mathcal{C} \exists ! \rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$. As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$. $\mathcal{L}(S,\mu) = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}. Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$. $\varrho$ is a homomorphism, as justified in \ref{classical_vector space}. $\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
\end{justification}

@@

Moreover, the density itself has to be differentiable. The distribution $\rho_\mathcal{c}$ is a real valued function of the state. As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$. As a state variable it identifies the states by the density of the material associated with them. This is more physically meaningful as one may first suspect, since placing physical markers is a common way to define the references for a coordinate system, which in the end is placing material at particular states. In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differential as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$. As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist. The density itself is differentiable.

\begin{prop}\label{differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a homogeneous classical material identified by a finite number of continuous independent state variables is a differentiable manifold. The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$. $\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$. Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

	We claim that the distributions associated with composite states are differentiable in the regions where they are strictly monotonic. Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$. Let $\rho_\mathcal{c}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$. Then its fibers in $U$, the inverse images of the values, are connected hyper-surfaces. We can construct a chart such that $q^1=\rho_\mathcal{c}$. The distribution is differentiable over $U$ for that variable: $\partial_{i} \rho_\mathcal{c} = \partial_{i} q^1 = \delta_i^1$ where $\delta_i^j$ is the Kronecker delta. Let $q^{j}$ be another set of state variables. $q^1(q^{j})$ is differentiable as the manifold is differentiable. $\rho_\mathcal{c}(q^{j})=q^1(q^{j})$ is differentiable over $U$.

	We claim that the distributions associated with composite states are differentiable. Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$. Let $\rho_{\mathcal{c}_0}$ be the distribution associated with state $\mathcal{c}_0$. Let $\rho_{\mathcal{c}_0}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$. Let $\delta \rho = \rho_{\mathcal{c}_0} - k \rho_{\mathcal{c}}$ where $k \in \mathbbm{R}$ and $k>0$. Let $k$ be sufficiently small so that $\delta \rho$ is strictly monotonic over $U$. $\rho_{\mathcal{c}_0}$ and $\delta \rho$ are differentiable over $U$ as they are strictly monotonic over $U$. $\rho_{\mathcal{c}}$ is a linear combination of differentiable functions over any arbitrary region $U$. $\rho_{\mathcal{c}}$ is differentiable.

	We claim $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$. Let $\rho_\mathcal{c}$ be the distribution associated with a state $\mathcal{c}$. Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function. If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$. Both integral are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$. Both integral are expressible with the same state variable as all elements are defined on the same differential structure. Therefore there exists an outcome for $\mathcal{S}$ that can physically distinguish the two distributions. There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho(\mathcal{c}_1)=\rho$. $\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective. $\varrho$ is a homomorphism, as justified in \ref{classical_vector space}. $\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
\end{justification}

As we identified the space of Lebesque integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill. Some are not restrictive enough. The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical. The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ not only include non-differentiable functions, but also include functions whose integral is infinite, which would represent infinite amount of material. Such functions are also present in the space of differentiable functions $C^1(\mathcal{S})$. Some definitions are too restrictive. Requiring compact support (i.e.the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states. Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range. Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible. The Whitney approximation theorem guarantees that this can be done mathematically.}

@@

If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary. This means we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This means that we cannot change physical units of all the variables as we like: they are part of a unit system. As we know, changing units of position will also change units of momentum. Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units. An element of $\mathcal{S}$ will be a geometrical object in such space. For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined. As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$. As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant. They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant. They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. Physically, the component $k_i$ of our covector, which are the classical analogue of the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$. If $q^i$ is in meters, $k_i$ is in inverse meters. The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation. A covector is the simplest local geometrical object that allows coordinate independent densities. The cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors, represents the simplest mathematical structure $\mathcal{S}$ can have to be able to support coordinate invariant densities.

If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$. These variables are independent, so the total number of possibilities in a small region of $\mathcal{S}$ is proportional to $dq \wedge dk$. That is if we double the range of $dq$ or $dk$ then the number of possibilities doubles. We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$. That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property. Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait. That is why one can choose "natural units" for which $\hbar=1$. In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e. compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$.

If $\mathcal{Q}$ is $n$-dimensional, $d\mu$ is proportional to $dq^n \wedge dk_n$. But it's not just the volume that is preserved as we change state variables. Each pair $(q^i, k_i)$ represents an independent set of state variables, with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$. We call such a pair a degree of freedom. As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$. This is equivalent to saying that independent d.o.f. are orthogonal surfaces in $\mathcal{S}$.

@@

\begin{justification}
	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form. The description of the simplest physical object on which we can define a state-variable-invariant must be given by at least one continuous state variable $q$. Let $\mathcal{Q}$ be the manifold charted by $q$. $\mathcal{Q}$ is an embedded manifold of $\mathcal{S}$. $\mathcal{Q}$ is differentiable. Let $q$ fully define the unit system for the description of the physical object. Then the physical object is represented by a geometrical object on $\mathcal{Q}$. The physical object is infinitesimal, therefore is represented by a local geometrical object. The simplest geometrical object in $\mathcal{Q}$ is a point, charted by $q$. A density over points is not invariant: $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$. Next we consider local geometrical objects charted by 2 variables. Densities over vectors $v(q) e_q \in T\mathcal{Q}$ are not invariant: $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$. Densities over covectors $k(q) dq \in T^*\mathcal{Q}$ are invariant: $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. Covectors are the simplest geometrical object to allow unit independent densities. The state space $\mathcal{S}$ for the simplest object that allows state-variable-independent densities is $T^*\mathcal{Q}$ of a one dimensional differential manifold $\mathcal{Q}$. Express integration in coordinates. $\int_{\mathcal{S}} \rho_\mathcal{c} d\mu \propto \int_{\mathcal{S}} \rho_\mathcal{c}(q, k) dq \wedge dk$. $d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant, $p\equiv \hbar k$  and $\omega$ is the canonical two form. $\omega$ is invariant under state variable change. $\omega$ is a symplectic metric for $\mathcal{S}=T^*\mathcal{Q}$. $\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
